BoostEM

### 2014-05-07 paper/paper.tex
algorithm thing
initial results written

\usepackage{algorithm}
\usepackage{algpseudocode}

\begin{document}
%
% --- Author Metadata here ---
\conferenceinfo{CS 760}{'14, Madison WI, USA}
\CopyrightYear{2014} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

@@

\section{Background}
\subsection{Boost}
\subsection{Expectation Maximization}
Expectation is an iterative algorithm which uses the values of existing,complete data points to compute the maximum likelihood estimate for missing values for other, incomplete data. Expectation Maximization was originally introduced in ~\cite{Dempster77maximumlikelihood} and produces a probabalistic distribution of the possible values of missing data given fully labeled data. In the context of semi-supervised learning, there is a set of labeled input data and a potentially much larger set of unlabeled data and for each unlabeled instance u, EM can provide a probability for each label.  These probabilities can be used as fractional instances of labeled data, making the input training set much bigger and diverse and potentially giving a better approximation of the distribution of labels to the learner. In many cases, performance can be drastically improved by using these fractional counts.

@@

\begin{algorithm}
\caption{BoostEM}\label{euclid}
\begin{algorithmic}[1]
\Procedure{BoostEM}{$L,U,T$}
\For{$t=1 \dots T$}
\State $P_{t} \gets EM(L,U,Wl,Wu)$
\EndFor
\State \textbf{return} $b$
\EndProcedure
\end{algorithmic}
\end{algorithm}

@@

@@

\section{Discussion}
STEVE 
\section{Conclusions}
%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}

>>>>

\usepackage{algorithm}
\usepackage{algpseudocode}
\begin{document}
%
% --- Author Metadata here ---
\conferenceinfo{CS 760}{'14, Madison WI, USA}
\CopyrightYear{2014} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

@@

\section{Background}
\subsection{Boost}

~\cite{Freund:1997:DGO:261540.261549}

\subsection{Expectation Maximization}
Expectation is an iterative algorithm which uses the values of existing,complete data points to compute the maximum likelihood estimate for missing values for other, incomplete data. Expectation Maximization was originally introduced in ~\cite{Dempster77maximumlikelihood} and produces a probabalistic distribution of the possible values of missing data given fully labeled data. In the context of semi-supervised learning, there is a set of labeled input data and a potentially much larger set of unlabeled data and for each unlabeled instance u, EM can provide a probability for each label.  These probabilities can be used as fractional instances of labeled data, making the input training set much bigger and diverse and potentially giving a better approximation of the distribution of labels to the learner. In many cases, performance can be drastically improved by using these fractional counts.

@@

\begin{algorithm}
\caption{BoostEM}\label{euclid}
\begin{algorithmic}[1]
\Procedure{BoostEM}{$L,U,H,T$}
\State L=Labeled Data
\State U=Unlabeled Data
\State H=Learner
\State T=Number of Boost iterations

\State for $i \in |L|$, $ Wl(i) \gets 1$
\State for $i \in |U|$, $ Wu(i) \gets 1$

\For{$t=1 \dots T$}
\State Normalize $Wu,Wl$
\State get fractional counts $P_{t} \gets EM(L,U,Wl,Wu)$
\State $H_{t} \gets \text{model learned from L, fractional counts of U}$
\State $\epsilon_{t} = \sum_{i=1}^{|L|} Wl_{i}(1-\delta(h_{t}(L_{i}),y_{i}))$
\State $\epsilon_{t} += \sum_{i=1}^{|U|} Wu_{i}(1-\delta(h_{t}(L_{i}),{P_{t,}}_{h_{t}(L_{i})}))$
\State if $|U| > 0$ then $\epsilon /= 2$
\State $\beta_{t} \gets \frac{\epsilon_{t}}{1-\epsilon_{t}}$
\State for $i,h_{t}(L_{i} = y_{i}) \in |L| Wl(i) \gets Wl(i)*\beta_{t}$
\State for $i \in |U|$, $Wl(i) \gets Wu(i)*\beta_{t}$

\EndFor
\State \textbf{return} $\arg \max_{y} \sum_{i=1}^{T}\bigg(\log(\frac{1}{\beta_{t}}\bigg) \delta(\bf{x},y)$
\EndProcedure
\end{algorithmic}
\end{algorithm}

@@

 Among the six algorithms we ran cross validation for, we expected FullBoost to have the best average performance, JustAlg to
have the worst average performance, and the other four algorithms to perform in between. In general this intuition was found to be true about FullBoost.
FullBoost was the top performing algorithm in 15 of the 16 settings we tested, with a setting being a dataset/labeled fraction/base learner triple. The only algorithm to best it was BoostEM on the parkinsons data set
with 20\% labeled data. JustAlg, however, only performed worse than all other algorithms in 5 settings. JustBoost and EM were each the worst performing algorithm
in 5 settings.

Ideally, BoostEM would in general perform better than both JustBoost and EM since it is a generalization and combination of the two techniques. This is the case, 
with BoostEM improving over both EM and Boost in 9 settings, and performing better than EM in 15 settings, and better than justBoost in 10. In no instances is BoostEM 
the worst performing of the three algorithms.

Somewhat surprisingly, BoostEM with 20\% of the data also outperformed using all the data on the base learner in 6 out of the 8 settings we tested.  This would seem to imply our learner can outperform a traditional base learner while only using one fifth of the labeled data.

BoostEM also performs well when the amount of labeled data is decreased by a large amount (one fourth).  In most cases, BoostEM retains its accuracy better than Boost and comparably to EM.

In general it seems that fluctuations in average accuracy when BoostEM uses a different base learner is mostly due to the base learner in question, as accuracy changes by BoostEM are reflected in accuracy changes using only the base learner.

@@

\section{Discussion}
STEVE 
YOU COULD TALK ABOUT HOW BECAUSE BOOSTEM CAN REPLICATE BOOST AND EM THAT WE ACTUALLY COULD DO REALLY WELL
\section{Conclusions}
%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}
